---
title: "Simulation of SIS model over networks"
author: "Marius Behret, Odysseas Kyparissis"
date: "`r Sys.Date()`"
output: html_document
---

# Loading libraries

```{r}
library(igraph)
set.seed(666) # For reproducibility
```

# Setting-up functions

```{r}
# Function to simulate SIS model on a given graph
sis_simulation <- function(g, beta, gamma, p0, time_steps) {
  set.seed(666)
  infected <- sample(c(0, 1), vcount(g), replace=TRUE, prob=c(1-p0, p0))
  status <- rep(0, vcount(g))
  results <- matrix(0, nrow=time_steps, ncol=vcount(g))

  for(t in 1:time_steps){
    for(i in 1:vcount(g)){
      if(infected[i] == 1){
        neighbors <- neighbors(g, i)
        for(j in neighbors){
          if(rbinom(1, size = 1, prob = beta)){
            status[j] <- 1
          }
        }
        if(rbinom(1, size = 1, prob = gamma)){
          status[i] <- 0
        } else {
          status[i] <- 1
        }
      }
    }
    infected <- status
    results[t, ] <- infected
  }
  rowSums(results) / vcount(g)
}

# Function to calculate the leading eigenvalue of a graph
leading_eigenvalue <- function(g) {
  eigen(as.matrix(get.adjacency(g)))[["values"]][1]
}

# Theoretical comparison of threshold
compare_numbers <- function(x, y) {
  if (x > y) {
    return(paste(x, "is greater than", y, "=>", "Epidemic"))
  } else if (x < y) {
    return(paste(x, "is less than", y, "=>", "NO Epidemic"))
  } else {
    return(paste(x, "is equal to", y))
  }
}

# Function to run simulations for a given network type
run_network_simulation <- function(network_type, n, m, k, p_er, p_ws, beta, gamma, p0, time_steps) {
  if (network_type == "ER") {
    g <- erdos.renyi.game(n, p_er, type="gnp")
  } else if (network_type == "BA") {
    g <- barabasi.game(n, m, directed=FALSE)
  } else if (network_type == "WS") {
    g <- watts.strogatz.game(1, n, k, p_ws, loops=FALSE, multiple=FALSE)
  } else if (network_type == "FC") {
    g <- make_full_graph(n)
  } else if (network_type == "Tree") {
    g <- make_tree(n)
  } else {
    stop("Invalid network type.")
  }

  # Simulate the SIS model
  infected_proportion <- sis_simulation(g, beta, gamma, p0, time_steps)

  # Calculate the leading eigenvalue
  eigen_val <- leading_eigenvalue(g)

  # Print relevant information
  cat("--------------------------------------------", "\n")
  cat("Network Type:", network_type, "\n")
  cat("Beta Divided by Gamma:", beta / gamma, "\n")
  cat("Leading eigenvalue:", eigen_val, "\n")
  threshold <- 1 / eigen_val
  cat("Threshold:", threshold, "\n")
  
  # Compare beta/gamma with the threshold
  cat(compare_numbers(beta / gamma, threshold), "\n")
  cat("--------------------------------------------", "\n")

  return(list(
    infected_proportion = infected_proportion,
    threshold = threshold,
    g = g
  ))
}


run_and_plot_sis_simulation <- function(n, m, k, p_er, p_ws, beta, gamma, p0, time_steps) {
  # Set up a larger plot
  par(mfrow=c(1, 1), mar=c(5, 5, 4, 2) + 0.1)
  
  # Plotting all infected_proportion variables together
  plot(1, type="n", main=paste0("SIS Model on Different Networks,", " beta=", beta, ", gamma=", gamma, ", p0=", p0), col="red", xlim = c(0, time_steps), ylim = c(0,1), xlab="Time", ylab="Infected Proportion")
  # Add subtitle
  subtitle = paste0(" m=", m, ", k=", k, ", p_er=", p_er, ", p_ws=", p_ws)
  title(sub = subtitle)
  
  # Networks to simulate
  network_types <- c("ER", "BA", "WS", "FC", "Tree")
  colors <- c("red", "blue", "darkgreen", "orange", "purple")
  
  # Store results in a list
  results_list <- list()
  
  # Simulate and plot for each network type
  for (i in seq_along(network_types)) {
    result <- run_network_simulation(network_types[i], n, m, k, p_er, p_ws, beta, gamma, p0, time_steps)
    
    lines(result$infected_proportion, col=colors[i])
    
    # Store result in the list
    results_list[[network_types[i]]] <- result
  }
  
  # Adding a legend outside the plot area
  legend("topright", legend=network_types, col=colors, lty=1, cex=0.5, inset=0.45)
  
  return(list(
    results_list = results_list,
    network_types = network_types
  ))
}

```

# Simulations of disease spread in different networks (Task 1)

To begin with, it is important to set-up proper values for the parameters of the networks and the spread of the infection. Thus:

-   n = 1000 (network size)
-   γ (gamma) = prob. that an infected node recovers
-   β (beta) = prob. that an infected node attempts to infect each neighbor
-   1/λ1 = epidemic threshold for arbitrary networks
-   p0 = initial random fraction of infected nodes

## Beta = 0.8, Gamma = 0.4: Epidemic

### m = 1, k_ws = 1, p_ws = 0.01, p_er = 0.01

```{r}
# Common Parameters
n <- 1000    # Number of nodes
beta <- 0.8  # Infection rate # try with 0.4 as well
gamma <- 0.4 # Recovery rate # and 0.8 for the results in the report
p0 <- 0.05   # Initial infection rate
time_steps <- 100 # Duration of the simulation
b.div.g <- beta / gamma
m <- 1 # m for barabasi.game: Number of edges to attach from a new node
k_ws <- 1 # k for watts.strogatz.game: Each node is connected to k nearest neighbors
p_ws <- 0.01 # p for watts.strogatz.game: Rewiring probability
p_er <- 0.01 # p for erdos.renyi.game: Probability for edge creation

results_epidemic1 <- run_and_plot_sis_simulation(n, m, k_ws, p_er, p_ws, beta, gamma, p0, time_steps)
```

### m = 10, k_ws = 10, p_ws = 0.1, p_er = 0.1

```{r}
# Common Parameters
n <- 1000    # Number of nodes
beta <- 0.8  # Infection rate # try with 0.4 as well
gamma <- 0.4 # Recovery rate # and 0.8 for the results in the report
p0 <- 0.05   # Initial infection rate
time_steps <- 100 # Duration of the simulation
b.div.g <- beta / gamma
m <- 10 # m for barabasi.game: Number of edges to attach from a new node
k_ws <- 10 # k for watts.strogatz.game: Each node is connected to k nearest neighbors
p_ws <- 0.1 # p for watts.strogatz.game: Rewiring probability
p_er <- 0.1 # p for erdos.renyi.game: Probability for edge creation

results_epidemic2 <- run_and_plot_sis_simulation(n, m, k_ws, p_er, p_ws, beta, gamma, p0, time_steps)
```



### m = 100, k_ws = 100, p_ws = 0.2, p_er = 0.2

```{r}
# Common Parameters
n <- 1000    # Number of nodes
beta <- 0.8  # Infection rate # try with 0.4 as well
gamma <- 0.4 # Recovery rate # and 0.8 for the results in the report
p0 <- 0.05   # Initial infection rate
time_steps <- 100 # Duration of the simulation
b.div.g <- beta / gamma
m <- 100 # m for barabasi.game: Number of edges to attach from a new node
k_ws <- 100 # k for watts.strogatz.game: Each node is connected to k nearest neighbors
p_ws <- 0.2 # p for watts.strogatz.game: Rewiring probability
p_er <- 0.2 # p for erdos.renyi.game: Probability for edge creation

results_epidemic3 <- run_and_plot_sis_simulation(n, m, k_ws, p_er, p_ws, beta, gamma, p0, time_steps)
```
## Beta = 0.4, Gamma = 0.8: NO Epidemic

### m = 1, k_ws = 1, p_ws = 0.01, p_er = 0.01

```{r}
# Common Parameters
n <- 1000    # Number of nodes
beta <- 0.4  # Infection rate # try with 0.4 as well
gamma <- 0.8 # Recovery rate # and 0.8 for the results in the report
p0 <- 0.05   # Initial infection rate
time_steps <- 100 # Duration of the simulation
b.div.g <- beta / gamma
m <- 1 # m for barabasi.game: Number of edges to attach from a new node
k_ws <- 1 # k for watts.strogatz.game: Each node is connected to k nearest neighbors
p_ws <- 0.01 # p for watts.strogatz.game: Rewiring probability
p_er <- 0.01 # p for erdos.renyi.game: Probability for edge creation

results_NO_epidemic1 <- run_and_plot_sis_simulation(n, m, k_ws, p_er, p_ws, beta, gamma, p0, time_steps)
```

### m = 10, k_ws = 10, p_ws = 0.1, p_er = 0.1

```{r}
# Common Parameters
n <- 1000    # Number of nodes
beta <- 0.4  # Infection rate # try with 0.4 as well
gamma <- 0.8 # Recovery rate # and 0.8 for the results in the report
p0 <- 0.05   # Initial infection rate
time_steps <- 100 # Duration of the simulation
b.div.g <- beta / gamma
m <- 10 # m for barabasi.game: Number of edges to attach from a new node
k_ws <- 10 # k for watts.strogatz.game: Each node is connected to k nearest neighbors
p_ws <- 0.1 # p for watts.strogatz.game: Rewiring probability
p_er <- 0.1 # p for erdos.renyi.game: Probability for edge creation

results_NO_epidemic2 <- run_and_plot_sis_simulation(n, m, k_ws, p_er, p_ws, beta, gamma, p0, time_steps)
```

### m = 100, k_ws = 100, p_ws = 0.2, p_er = 0.2

```{r}
# Common Parameters
n <- 1000    # Number of nodes
beta <- 0.4  # Infection rate # try with 0.4 as well
gamma <- 0.8 # Recovery rate # and 0.8 for the results in the report
p0 <- 0.05   # Initial infection rate
time_steps <- 100 # Duration of the simulation
b.div.g <- beta / gamma
m <- 100 # m for barabasi.game: Number of edges to attach from a new node
k_ws <- 100 # k for watts.strogatz.game: Each node is connected to k nearest neighbors
p_ws <- 0.2 # p for watts.strogatz.game: Rewiring probability
p_er <- 0.2 # p for erdos.renyi.game: Probability for edge creation

results_NO_epidemic3 <- run_and_plot_sis_simulation(n, m, k_ws, p_er, p_ws, beta, gamma, p0, time_steps)
```

# Simulations of disease spread based on threshold value in different networks (Task 2)

```{r}
# Function to run simulations for two sets of parameter values
run_threshold_simulations <- function(g, p0, time_steps, threshold, epsilon = 0.05) {
  
  if (threshold == Inf) {
    beta_below <- 0.001
    gamma_below <- 0.1
    beta_above <- 0.1
    gamma_above <- 0.001
    threshold_below = Inf
    threshold_above = Inf
  } else {
    threshold_below = threshold - epsilon
    if (threshold_below < 0){
      threshold_below = threshold - 0.0002
    }
    res = generate_random_values(threshold_below)
    beta_below <- res[1]
    gamma_below <- res[2]
    threshold_above = threshold + epsilon
    res = generate_random_values(threshold_above)
    beta_above <- res[1]
    gamma_above <- res[2]
  }
  

  # Simulate below threshold
  infected_proportion_below <- sis_simulation(g, beta_below, gamma_below, p0, time_steps)

  # Simulate above threshold
  infected_proportion_above <- sis_simulation(g, beta_above, gamma_above, p0, time_steps)

  list(
    infected_proportion_below = infected_proportion_below,
    infected_proportion_above = infected_proportion_above,
    threshold_below = threshold_below,
    beta_below,
    gamma_below,
    threshold_above = threshold_above,
    beta_above = beta_above,
    gamma_above = gamma_above
  )
}


generate_random_values <- function(threshold) {
  # Generate a random value for x within [0, 1]
  gamma <- runif(1, min = 0.001, max = 1)
  
  beta <- gamma * threshold
  
  # Return the values
  return(c(beta = beta, gamma = gamma))
}


get_threshold_simulations <- function(results, p0, time_steps) {
  
  
  # Networks to simulate
  network_types <- c("ER", "BA", "WS", "FC", "Tree")
  colors <- c("red", "blue", "darkgreen", "orange", "purple")
  
  # Store results in a list
  results_list <- list()
  
  # Simulate and plot for each network type
  for (i in seq_along(network_types)) {
    if (network_types[i] == "ER") {
      data <- results$results_list[i]$ER
    } else if (network_types[i] == "BA") {
      data <- results$results_list[i]$BA
    } else if (network_types[i] == "WS") {
      data <- results$results_list[i]$WS 
    } else if (network_types[i] == "FC") {
      data <- results$results_list[i]$FC
    } else if (network_types[i] == "Tree") {
      data <- results$results_list[i]$Tree
    } else {
      stop("Invalid network type.")
    }
    
    result <- run_threshold_simulations(data$g, p0, time_steps, data$threshold)
    
    # Store result in the list
    results_list[[network_types[i]]] <- result
  }
  
  return(list(
    results_list = results_list,
    network_types = network_types
  ))
}

```

```{r}
# Threshold Simulations
task2_epidemic1 <- get_threshold_simulations(results_epidemic1, p0, time_steps)
task2_epidemic2 <- get_threshold_simulations(results_epidemic2, p0, time_steps)
task2_epidemic3 <- get_threshold_simulations(results_epidemic3, p0, time_steps)
# task2_NOepidemic1 <- get_threshold_simulations(results_NO_epidemic1, p0, time_steps)
# task2_NOepidemic2 <- get_threshold_simulations(results_NO_epidemic2, p0, time_steps)
# task2_NOepidemic3 <- get_threshold_simulations(results_NO_epidemic3, p0, time_steps)
```


## Erdős-Rényi network


```{r}
# Plotting the results for Erdős-Rényi network
colors <- c("red", "blue", "darkgreen", "orange", "purple")
par(mfrow=c(1, 1))
plot(task2_epidemic1$results_list$ER$infected_proportion_below, type="l", col="blue", ylim=c(0, 1.3), main="Threshold Check (ER)", ylab = "Infected Proportion", lty =1)
lines(task2_epidemic1$results_list$ER$infected_proportion_above, col="darkgreen", lty =2)
lines(task2_epidemic2$results_list$ER$infected_proportion_below, col="orange", lty =1)
lines(task2_epidemic2$results_list$ER$infected_proportion_above, col="purple", lty =2)
lines(task2_epidemic3$results_list$ER$infected_proportion_below, col="red", lty =1)
lines(task2_epidemic3$results_list$ER$infected_proportion_above, col="lightblue", lty =2)

legend("topright", legend=c("Below, m = 1, k_ws = 1, p_ws = 0.01, p_er = 0.01", "Above, m = 1, k_ws = 1, p_ws = 0.01, p_er = 0.01", "Below, m = 10, k_ws = 10, p_ws = 0.1, p_er = 0.1", "Above, m = 10, k_ws = 10, p_ws = 0.1, p_er = 0.1", "Below, m = 100, k_ws = 100, p_ws = 0.2, p_er = 0.2", "Above, m = 100, k_ws = 100, p_ws = 0.2, p_er = 0.2"), col=c("blue", "darkgreen","orange","purple", "red", "lightblue"), lty=c(1,2,1,2,1,2), cex=0.5)

```

## Barabási-Albert network

```{r}
# Plotting the results for Erdős-Rényi network
colors <- c("red", "blue", "darkgreen", "orange", "purple")
par(mfrow=c(1, 1))
plot(task2_epidemic1$results_list$BA$infected_proportion_below, type="l", col="blue", ylim=c(0, 0.25), main="Threshold Check (BA)", ylab = "Infected Proportion", lty =1)
lines(task2_epidemic1$results_list$BA$infected_proportion_above, col="darkgreen", lty =2)
lines(task2_epidemic2$results_list$BA$infected_proportion_below, col="orange", lty =1)
lines(task2_epidemic2$results_list$BA$infected_proportion_above, col="purple", lty =2)
lines(task2_epidemic3$results_list$BA$infected_proportion_below, col="red", lty =1)
lines(task2_epidemic3$results_list$BA$infected_proportion_above, col="lightblue", lty =2)

legend("topright", legend=c("Below, m = 1, k_ws = 1, p_ws = 0.01, p_er = 0.01", "Above, m = 1, k_ws = 1, p_ws = 0.01, p_er = 0.01", "Below, m = 10, k_ws = 10, p_ws = 0.1, p_er = 0.1", "Above, m = 10, k_ws = 10, p_ws = 0.1, p_er = 0.1", "Below, m = 100, k_ws = 100, p_ws = 0.2, p_er = 0.2", "Above, m = 100, k_ws = 100, p_ws = 0.2, p_er = 0.2"), col=c("blue", "darkgreen","orange","purple", "red", "lightblue"), lty=c(1,2,1,2,1,2), cex=0.5)
```

## Watts-Strogatz network

```{r}
# Plotting the results for Erdős-Rényi network
colors <- c("red", "blue", "darkgreen", "orange", "purple")
par(mfrow=c(1, 1))
plot(task2_epidemic1$results_list$WS$infected_proportion_below, type="l", col="blue", ylim=c(0, 1.3), main="Threshold Check (WS)", ylab = "Infected Proportion", lty =1)
lines(task2_epidemic1$results_list$WS$infected_proportion_above, col="darkgreen", lty =2)
lines(task2_epidemic2$results_list$WS$infected_proportion_below, col="orange", lty =1)
lines(task2_epidemic2$results_list$WS$infected_proportion_above, col="purple", lty =2)
lines(task2_epidemic3$results_list$WS$infected_proportion_below, col="red", lty =1)
lines(task2_epidemic3$results_list$WS$infected_proportion_above, col="lightblue", lty =2)

legend("topright", legend=c("Below, m = 1, k_ws = 1, p_ws = 0.01, p_er = 0.01", "Above, m = 1, k_ws = 1, p_ws = 0.01, p_er = 0.01", "Below, m = 10, k_ws = 10, p_ws = 0.1, p_er = 0.1", "Above, m = 10, k_ws = 10, p_ws = 0.1, p_er = 0.1", "Below, m = 100, k_ws = 100, p_ws = 0.2, p_er = 0.2", "Above, m = 100, k_ws = 100, p_ws = 0.2, p_er = 0.2"), col=c("blue", "darkgreen","orange","purple", "red", "lightblue"), lty=c(1,2,1,2,1,2), cex=0.5)
```
## Tree network

```{r}
# Plotting the results for Erdős-Rényi network
colors <- c("red", "blue", "darkgreen", "orange", "purple")
par(mfrow=c(1, 1))
plot(task2_epidemic1$results_list$Tree$infected_proportion_below, type="l", col="blue", ylim=c(0, 0.9), main="Threshold Check (Tree)", ylab = "Infected Proportion", lty =1)
lines(task2_epidemic1$results_list$Tree$infected_proportion_above, col="darkgreen", lty =2)
lines(task2_epidemic2$results_list$Tree$infected_proportion_below, col="orange", lty =1)
lines(task2_epidemic2$results_list$Tree$infected_proportion_above, col="purple", lty =2)
lines(task2_epidemic3$results_list$Tree$infected_proportion_below, col="red", lty =1)
lines(task2_epidemic3$results_list$Tree$infected_proportion_above, col="lightblue", lty =2)

legend("topright", legend=c("Below, m = 1, k_ws = 1, p_ws = 0.01, p_er = 0.01", "Above, m = 1, k_ws = 1, p_ws = 0.01, p_er = 0.01", "Below, m = 10, k_ws = 10, p_ws = 0.1, p_er = 0.1", "Above, m = 10, k_ws = 10, p_ws = 0.1, p_er = 0.1", "Below, m = 100, k_ws = 100, p_ws = 0.2, p_er = 0.2", "Above, m = 100, k_ws = 100, p_ws = 0.2, p_er = 0.2"), col=c("blue", "darkgreen","orange","purple", "red", "lightblue"), lty=c(1,2,1,2,1,2), cex=0.5)
```
## Fully Connected


```{r}
# Plotting the results for Fully Connected network
colors <- c("red", "blue", "darkgreen", "orange", "purple")
par(mfrow=c(1, 1))
plot(task2_epidemic1$results_list$FC$infected_proportion_below, type="l", col="blue", ylim=c(0, 1.4), main="Threshold Check (FC)", ylab = "Infected Proportion", lty =1)
lines(task2_epidemic1$results_list$FC$infected_proportion_above, col="darkgreen", lty =2)
lines(task2_epidemic2$results_list$FC$infected_proportion_below, col="orange", lty =1)
lines(task2_epidemic2$results_list$FC$infected_proportion_above, col="purple", lty =2)
lines(task2_epidemic3$results_list$FC$infected_proportion_below, col="red", lty =1)
lines(task2_epidemic3$results_list$FC$infected_proportion_above, col="lightblue", lty =2)

legend("topright", legend=c("Below, m = 1, k_ws = 1, p_ws = 0.01, p_er = 0.01", "Above, m = 1, k_ws = 1, p_ws = 0.01, p_er = 0.01", "Below, m = 10, k_ws = 10, p_ws = 0.1, p_er = 0.1", "Above, m = 10, k_ws = 10, p_ws = 0.1, p_er = 0.1", "Below, m = 100, k_ws = 100, p_ws = 0.2, p_er = 0.2", "Above, m = 100, k_ws = 100, p_ws = 0.2, p_er = 0.2"), col=c("blue", "darkgreen","orange","purple", "red", "lightblue"), lty=c(1,2,1,2,1,2), cex=0.5)
```

