---
title: "Finding and Assessing Community Structure"
author: "Liam Glennie, Odysseas Kyparissis"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Importing Libraries

```{r}
library(igraph)
library(clustAnalytics)
library(sets)
```

# Importing, Generating and Preparing Graph Data (karate, synthetic graph, enron, rfid)

```{r, warning = FALSE}

# karate
data(karate,package="igraphdata");
karate <- upgrade_graph(karate);

# a synthetic network with scale-free degree distribution (200 nodes, 800 edges and 4 communities)
B = matrix(c(1, 0.25, 0.25, 0.25, 0.25, 1, 0.25, 0.25, 0.25, 0.25, 1, 0.25, 0.25, 0.25, 0.25, 1), ncol = 4)
synthetic_network = barabasi_albert_blocks(m = 4, p = c(0.25, 0.25, 0.25, 0.25), B = B, t_max = 200, type = "Hajek", sample_with_replacement = FALSE)

# enron
data(enron,package="igraphdata");
enron <- upgrade_graph(enron);

## Convert the adjacency matrix to an igraph object
enron <- graph.adjacency(as_adjacency_matrix(as.undirected(enron, mode = "each")), mode = "undirected", weighted = TRUE, diag = FALSE)

# rfid
data(rfid,package="igraphdata");
rfid <- upgrade_graph(rfid);
```

# Tasks
Below the 4 requested tasks are being developed.

## Task 1: Jaccard Index Computation (jaccard_sim)

```{r}
jaccard_sim <- function(members_a, members_b) {
  row_names <- unique(members_a)
  column_names <- unique(members_b)
  
  jaccard_table <- data.frame(matrix(nrow = length(row_names), ncol =  length(column_names)))
  rownames(jaccard_table) <- row_names
  colnames(jaccard_table) <- column_names

  for (i in row_names) {
    ids_a <- which(members_a==i)
    for (j in column_names) {
      ids_b <- which(members_b==j)
      jaccard_table[i, j] <- length(intersect(ids_a, ids_b)) / length(union(ids_a, ids_b))
    }
  }
  return(jaccard_table)
}
```

## Task 2: Cluster Similarity Identification (match_clusters)


```{r}
match_clusters <- function(jaccard_table, name_1, name_2) {
  result_table <- data.frame(matrix(nrow = 1, ncol = nrow(jaccard_table)))
  
  for (a in 1:nrow(jaccard_table)) {
    max_idx <- which(jaccard_table[a,] == max(jaccard_table[a,]))
    result_table[1,a] <- jaccard_table[a,max_idx]
    names(result_table)[a] <- paste0('(', name_1, '.', a, ',', name_2, '.', max_idx, ')')
  }
  
  return(result_table)
}
```

## Task 3: Weighted Mean Calculation for Global Jaccard Similarity (Wmean)

```{r}
Wmean <- function(mc) {
  # Calculate the average of Jaccard indices
  
  # Convert the first row to numeric
  numeric <- as.numeric(as.character(mc[1, ]))
  
  # Calculate the mean
  globalJaccardIndex <- mean(numeric, na.rm = TRUE)
  
  return(globalJaccardIndex)
}
```

## Task 4: Evaluation of Community Detection Algorithms

```{r}
# Function to evaluate significance of clusterings using selected scoring functions
evaluate_clusterings <- function(graph, algorithms, ground_truth = NULL) {

  # Evaluate significance using the clustAnalytics package
  significance_table <- evaluate_significance(
    graph,
    alg_list = algorithms,
    gt_clustering = ground_truth
  )

  return(significance_table)
}
```

Here, the lists of algorithms and networks to be analyzed are established for the experiment.

```{r}
# List of community detection algorithms to evaluate
algorithms <- list(
  Louvain = cluster_louvain,
  "Label Propagation" = cluster_label_prop,
  Walktrap = cluster_walktrap,
  betweenness=cluster_edge_betweenness
)

# Networks to evaluate
networks <- list(
  karate = karate,
  synthetic = synthetic_network,
  enron = enron,
  rfid = rfid
)
```


```{r}
for (network_name in names(networks)) {
  print(network_name)
  network <- networks[[network_name]]
  
  # Evaluate clusterings using the specified algorithms
  significance_table <- evaluate_clusterings(network, algorithms)
  
  # select a representative metric of each class
  # Internal Connectivity. Internal Density
  # External Connectivity. Cut Ratio
  # Combine Internal and External Connectivity. Normalised Cut
  # Based on a Network Model. Modularity
  selected_metrics <- significance_table[c("internal density", "cut ratio",
                                        "norm cut", "modularity"), ]
  
  print(selected_metrics)
  
  
  # Find a good reference clustering (ground truth)
  
  # If the network has a known ground truth
  # set it as the ground truth clustering
  # Otherwise calculate a ground truth
  # based on the selective metrics
  # and an average mean between them
  # by giving negative values to metrics
  # which is better to have small values
  
  ground_truth <- NULL
  if (network_name ==  "karate") {
    
    ground_truth <- V(network)$Faction
  
  } else if (network_name ==  "synthetic"){
    
    ground_truth <- V(network)$label
  
  } else{
    
    # Assuming selected_metrics is your data frame
    selected_metrics <- as.matrix(selected_metrics)
    
    # Define weights (you can adjust these according to your needs)
    weights <- c(1, -1, -1, 1)
    
    # Calculate weighted mean for each column
    weighted_means <- colSums(selected_metrics * weights) / sum(abs(weights))
    
    # Create a new data frame with the results
    result_table <- data.frame(
      Weighted_Mean = weighted_means,
      row.names = colnames(selected_metrics)
    )
    
    print("Ground truth for:")
    print(network_name)
    print("is:")
    print(rownames(result_table)[which.max(result_table$Weighted_Mean)])
    
    best_algorithm <- algorithms[[rownames(result_table)[which.max(result_table$Weighted_Mean)]]]
    ground_truth <- unname(membership(best_algorithm(network)))
  
  }
  
  # Evaluation of clustering with functions from Tasks 1-3
  
  for (algorithm_name in names(algorithms)) {
    print(algorithm_name)
    algorithm <- algorithms[[algorithm_name]]
    
    # Run the algorithm on the network
    membership <- algorithm(network)$membership
    
    # Compute Jaccard similarity
    jaccard_table <- jaccard_sim(ground_truth, membership)
    
    # Match clusters and calculate weighted mean
    matched_clusters <- match_clusters(jaccard_table, name_1 = "GT", name_2 = algorithm_name)
    global_jaccard_similarity <- Wmean(matched_clusters)
    
    # Print the results for each algorithm on the current network
    cat(paste("Results for", algorithm_name, "algorithm on", network_name, "network:\n"))
    print(jaccard_table)
    print(matched_clusters)
    cat("Global Jaccard Similarity:", global_jaccard_similarity, "\n\n")
    
  }
}
```

